---
title: "Candidate Variants"
output: html_notebook
author: "Diana Barrientos González"
date: "31-05-2026"
---


## Database exploration

### Candidate variants for autosomal recessive disorder

```{r}
cat("Genes with candidate variants of autosomal recessive disorder:\n")
table(filtro10[,19])
```

- BAIAP2L2 - **DISCARDED**

There is no information in ClinVar but it is classified as STOP GAINED, we see that there is a phenotype in OMIM of this gene and in its band but it does not coincide with our phenotype.

- C16orf96  - NM_001145011 - **DISCARDED**

    - NCBI: C16orf96:C16orf96:NM_001145011:exon9:c.G2578A:p.D860N,

    - Variable identifier: NM_001145011.2(C16orf96):c.2578G>A (p.Asp860Asn)

    - dbSNP: rs150345449

    - Category on Clinvar: **Likely benign**

- C3orf52 - **DISCARDED**

    - c.397-5->T as it is from splicing and is not in a canonical base we can leave it as HALF discarded, in Clinvar it has no clinical significance.

- CTBS- NM_004388 - **DISCARDEDA**

    - NCBI: CTBS:NM_004388:exon6:c.G820A:p.V274I,

    - dbSNP : 	rs142534762

Two other rsIDs have been merged into rs3217269, but it is not reported in ClinVar.

For this variant, we can go to Uniprot and check if the deletion occurs in a domain. If not, it is unlikely to be the cause (p.31_34del), and it is also a non-frameshift deletion.

The domain reported in UNIPROT is in positions 39-385. It does not affect a functional site, so it is unlikely to be the cause, but we should continue investigating.

Let's check OMIM to see if there are any reported clinical phenotypes. There are two phenotypes in that band, none in exon 1 or at those nucleotides, and no related phenotype appears.


- ORC4 - **DISCARDED**

    - We checked dbSNP from NCBI and found that we can search by rsID: rs2307399

    - It says it has: Clinical significance: **benign**

- PAX7 - **CANDIDATE**

It is not found in Clinvar, we checked in OMIM and it appears as: #618578 CONGENITAL MYOPATHY 19; CMYO19, whioch matches with the given phenotype

- PTDSS1 (splicing) - **DISCARDED**

    - There is a phenotype in OMIM, but it corresponds to a heterozygous mutation and the phenotype doesn't match; besides, the location is not the same. We ruled it out.

- RTKN2 **HALF DISCARDED**

    - It's not found in Clinvar, we checked OMIM and it's not related

- SLC3A1 **DISCARDED**

    - It is not found in clinvar, and was ruled out in OMIM.

- TRAK2 **DISCARDED**

    - Classified as Benign and merged into rs3832088

- TRIOBP **HALF DISCARDED**

    - It has a similar phenotype, but deafness on it, we rule it out

- ZNF490 **HALF DISCARDED**

    - Uncertain significance, no clinical feature

### Genes with candidate variants of compound heterozygosity

```{r}
cat("Genes with candidate variants of compound heterozygosity:\n")
table(filtro15[,19])
```

As we already have a likely disease, we don't need to check one by one again, we can just look at OMIM

- CARD8 : It does not have a phenotype in OMIM that is consistent with the given one.

- CCDC168:  Not found in OMIM

- COL11A1:  #228520 FIBROCHONDROGENESIS 1; FBCG1, #154780 MARSHALL SYNDROME; MRSHS & #604841 STICKLER SYNDROME, TYPE II; STL2

- COL5A1: OMIM states that the mutation is heterozygous (not compound heterozygous).

- CRISPLD2: Nothing on OMIM

- DOCK6: There is compound heterozygosity, but it doesn't fit

- DYSF: #254130 MIYOSHI MUSCULAR DYSTROPHY 1; MMD1, 

- FCGBP: No phenotype

- FGFR3: Yes, phenotypes, but not due to compound heterozygosity

- GPR182: No clinical phenotype

- GPR98: Doesn't match our clinical phenotype

- NUTM2B:  Doesn't match our clinical phenotype

- PRRC2A: No clinical phenotype

- ROBO3: No clinical phenotype

- SLC39A14: The phenotype does not have compound heterozygosity

- STX2: No phenotype

- TNC:here is a phenotype but no compound heterozygosity

- TRIM73: No ohenotype


## FINAL CANDIDATE

The described phenotype —global central hypotonia, gross motor delay, and mild facial dysmorphia— constitutes the cardinal features of Congenital Myopathy due to *PAX7* (CMYO19). The presence of transient torticollis is also consistent with the cervical muscle weakness described in this condition.

The other phenotypes seen did not match as much as this one.